Under-approximation of reachability through static analysis in multivalued asynchronous networks

### 2013-03-28 parts/def-ph.tex
demo

Given this first hypothesis, we define the neighbor \towrite{revoir terminologie : neighbors ? vicinity ?} sorts and actions (\pref{def:neighbors}).
\begin{definition}
\label{def:chainsactions}
  We define the set of chains of actions $\PHh(b, a)$ between two sorts $b$ and $a$ as below:
  \begin{align*}
  \PHh(b, a) = \{ (h_i)_{i \in \segm{0}{s}} \in (\PHh^{(1)})^{s+1} &\mid s \in \sN, \PHsort(\PHhitter(h_0)) = b \wedge \PHsort(\PHtarget(h_s)) = a \\
    & \wedge \forall i \in \segm{1}{s}, \PHsort(\PHtarget(h_{i-1})) = \PHsort(\PHhitter(h_{i})) \}
  \end{align*}
\end{definition}
\begin{definition}
\label{def:neighbors}
  The set of neighbor sorts $\Vs(a)$ of a given sort $a$ is the set of sorts that can interact with it through actions of priority $1$.
  The set of neighbor actions $\Vh(a)$ of $a$ is the set of actions of priority $1$ influencing $a$.
  $$\Vs(a) = \{ b \in \PHs \mid \PHh(b, a) \neq \emptyset \}$$
  $$\Vh(a) = \{ h \in \PHh^{(1)} \mid \PHsort(\PHhitter(h)) \in \Vs(a) \wedge \PHsort(\PHtarget(h)) \in \Vs(a) \cup \{ a \} \}$$
\end{definition}

In \pref{def:cs} and \pref{def:component} we define the notion of well-formed component and cooperative sort.
A sort $a$ is a well-formed component iff
it is hit only by actions of priority $2$.
A cooperative sort $a$ is well-formed iff
it is hit only by actions of priority $1$,
if any sub-state of its neighborhood is represented by at least one local steady state in this sort,
and if any of its processes represents at least one sub-state of its neighborhood.
\begin{definition}
\label{def:component}
  A sort $a \in \PHs$ is said to be a well-formed component iff:
  \begin{itemize}
    \item $\forall h \in \PHh, \PHsort(\PHtarget(h)) = a \Rightarrow \prio(h) = 2$
  \end{itemize}
\end{definition}
\begin{definition}
  \label{def:cs}
  A sort $a \in \PHs$ is said to be a well-formed cooperative sort iff:
  \begin{itemize}
    \item $\forall h \in \PHh, \PHsort(\PHtarget(h)) = a \Rightarrow \prio(h) = 1$
    \item $\forall s \in \underset{b \in \Vs(a)}{\times} \PHl_b, \pfp_s(a) \neq \emptyset$
    \item $\forall a_i \in \PHl_a, \exists s \in \underset{b \in \Vs(a)}{\times} \PHl_b, a_i \in \pfp_s(a)$
  \end{itemize}
\end{definition}
We call $\components$ the set of well-formed components and $\cs$ the set of well-formed cooperative sorts.
We note that: $\forall a \in \components, \Vs(a) = \emptyset \wedge \Vh(a) = \emptyset$

Finally, we denote $\pfp_s(a)$ the set of possible local steady-states of a sort $a$ from a state $s$.
États focaux possibles: pour tout contexte $\ctx \in \restriction{\PHl}{\Vs(a)}$:
\begin{align*}
  \pfp_\ctx(a) = \{ a_i \in \PHl_a &\mid \exists s \in \ctx, \exists \delta \in \restriction{\Sce}{1}, \PHget{(\ctx \PHplay \delta)}{a} = a_i \\
  & \quad \wedge \forall \PHhit{b_i}{c_j}{c_k} \in \Vh(a), \PHget{(s \PHplay \delta)}{b} \neq b_i \vee \PHget{(s \PHplay \delta)}{c} \neq c_j \}
\end{align*}
Of couse, if $a \in \components$, then $\pfp_\ctx(a) = \PHget{\ctx}{a}$.

>>>>

In \pref{def:cs} we define the notion of well-formed component.
A sort $a$ is a well-formed component iff
it is hit only by actions of priority $2$.
\begin{definition}
\label{def:component}
  A sort $a \in \PHs$ is said to be a well-formed component iff:
  \begin{itemize}
    \item $\forall h \in \PHh, \PHsort(\PHtarget(h)) = a \Rightarrow \prio(h) = 2$
  \end{itemize}
\end{definition}
We call $\components$ the set of well-formed components.

Given this first hypothesis, we define the neighbor \towrite{revoir terminologie : neighbors ? vicinity ?} sorts and actions (\pref{def:neighbors}).
\begin{definition}
\label{def:chainsactions}
  We define the set of chains of actions $\PHh(b, a)$ between two sorts $b$ and $a$ as below:
  \begin{align*}
  \PHh(b, a) = \{ (h_i)_{i \in \segm{0}{s}} \in (\PHh^{(1)})^{s+1} &\mid s \in \sN, \PHsort(\PHhitter(h_0)) = b \wedge \PHsort(\PHtarget(h_s)) = a \\
    & \wedge \forall i \in \segm{1}{s}, \PHsort(\PHtarget(h_{i-1})) = \PHsort(\PHhitter(h_{i})) \}
  \end{align*}
\end{definition}
\begin{definition}
\label{def:neighbors}
  The set of neighbour sorts $\Vs(a)$ of a given sort $a$ is the set of components that can interact with it through actions of priority $1$.
  The set of neighbour actions $\Vh(a)$ of $a$ is the set of actions of priority $1$ influencing $a$.
  $$\Vs(a) = \{ b \in \components \mid \PHh(b, a) \neq \emptyset \}$$
  $$\Vh(a) = \{ h \in \PHh^{(1)} \mid \PHsort(\PHhitter(h)) \in \Vs(a) \wedge \PHsort(\PHtarget(h)) \in \Vs(a) \cup \{ a \} \}$$
\end{definition}

We denote $\pfp_s(a)$ the set of possible local steady-states of a sort $a$ from a state $s$.
\todo{À corriger : état global ? + prendre en compte l'état de la sorte}
\begin{align*}
  \pfp_s(a) = \{ a_i \in \PHl_a &\mid \exists \delta \in \restriction{\Sce}{1}, \PHget{(s \PHplay \delta)}{a} = a_i \\
  & \quad \wedge \forall \PHhit{b_i}{c_j}{c_k} \in \Vh(a), \PHget{(s \PHplay \delta)}{b} \neq b_i \vee \PHget{(s \PHplay \delta)}{c} \neq c_j \}
\end{align*}
Of course, if $a \in \components$, then $\pfp_s(a) = \PHget{s}{a}$.

In \pref{def:component} we define the notion of well-formed cooperative sort.
A cooperative sort $a$ is well-formed iff
it is hit only by actions of priority $1$,
if any sub-state of its neighbourhood is represented by at least one local steady state in this sort,
and if any of its processes represents at least one sub-state of its neighbourhood.
\begin{definition}
\label{def:cs}
  A sort $a \in \PHs$ is said to be a well-formed cooperative sort iff:
  \begin{itemize}
    \item $\forall h \in \PHh, \PHsort(\PHtarget(h)) = a \Rightarrow \prio(h) = 1$
    \item $\forall s \in \underset{b \in \Vs(a)}{\times} \PHl_b, |\pfp_s(a)| = 1$
    \item $\forall a_i \in \PHl_a, \exists s \in \underset{b \in \Vs(a)}{\times} \PHl_b, a_i \in \pfp_s(a)$
  \end{itemize}
\end{definition}
We call $\cs$ the set of well-formed cooperative sorts.
%We note that: $\forall a \in \components, \Vs(a) = \emptyset \wedge \Vh(a) = \emptyset$
As any local state of neighbour sorts of a cooperative sort leads to a local steady-state, we denote $\csState(a_i)$ the set of local states represented by the process $a_i$ of the cooperative sort $a$.
\begin{definition}[$\csState$]
\label{def:csState}
  \towrite{À reformuler}
  $$\csState(a_i) = \{ s \in \underset{b \in \Vs(a)}{\times} \PHl_b \mid \pfp_{s}(a) = a_i \}$$
\end{definition}

### 2013-03-28 parts/sa.tex
demo

\begin{definition}[$\csState$]
  \label{def:csState}
  \towrite{À reformuler}
  $$\csState(a_i) = \{ b_j \in \PHproc \mid b \in \Vs(a) \wedge \pfp_{\csState(a_i)}(a) = a_i \}$$
\end{definition}

\begin{definition}
  \label{def:aS}
  The abstract structure $\cwB=(\Breq,\Bsol,\Brsp,\Bcont)$ is defined as
  $\cwB = \sfp{\aB^\w_\ctx}{\myB}{\aB^\w_{\ctx \Cap \allprocs_\ctx(\myB)}}$,\\
  with $\myB=(\myreq,\mysol,\myrsp,\mycont)$:
  \begin{align*}
    \myreq &= \{ (a_i,\PHobjp{a}{j}{i}) \in \PHproc \times \Obj \mid
      a \in \components \wedge a_j \in \PHget{\ctx}{a} \\ % \vee a_j \in \pfpprocs_\ctx(a) \\
      & \qquad \wedge ((\exists (P,ps) \in \mysol \vee \exists (b_j, ps) \in \myrsp), a_i \in ps \\
      & \qquad\qquad \vee \exists n \in \indexes{\w}, \PHbounce(\w_n)=a_i) \}
    \\
    \mysol &\subseteq \{ (P,ps) \in \Obj \times \powerset(\PHproc) \mid
            \exists (a_i, P) \in \myreq \wedge ps \in \aBS(P) \\
      & \qquad\qquad \vee \exists (Q, P) \in \mycont \wedge ps \in \aBS(P) \}
    \\
    \myrsp &= \{ (a_i,ps) \in \PHproc \times \powerset(\PHproc) \mid a \in \cs \\
      & \qquad \wedge (\exists (P,ps') \in \mysol \vee \exists (b_j,ps') \in \myrsp), \\
      & \qquad\qquad a_i \in ps' \wedge \csState(a_i) = \towrite{\in?} ps \}
    \\
    \mycont & = \{ (P, \PHobj{q}{\PHbounce(P)}) \in \Obj \times \Obj \mid
      \exists (P, ps) \in \mysol \\
      & \qquad\qquad \wedge q \in \gCont_\ctx(\PHsort(P),P) \}
  \end{align*}
\end{definition}

\begin{definition}[Coherent solution]\label{def:coherent}
  A cooperative sort requisite $(a_i, ps)$ in $\myrsp$ is said coherent iff
  it has no successor $(b_k,\PHobjp{b}{j}{k})$ in $\myreq$ so that there exists $b_n \in ps$, $b_k \neq b_n$.
\end{definition}

\begin{theorem}[Approximation inf.]\label{th:approxinf}
  If the graph $\cwB$ contains no cycle,
  all objectives have at least one solution
  and all cooperative sort requisites are coherent,
  then $\uconcr(\w) \neq \emptyset$.
\end{theorem}

@@

\towrite{MQ par récurrence : tous les $a_i$ requis de sortes coopératives sont concrétisables}

@@

  \item Suppose all children objectives of $P$ are concretizable.
  \begin{itemize}
    \item If $\exists (P, Q) \in \Bcont$, then by hypothesis,
      $\muconcr_{s}(\obj{\PHtarget(P)}{\PHtarget(Q)} \concat Q) \neq \emptyset$, thus
      $\muconcr_{s}(P) \neq \emptyset$.
    \item Else, by \pref{def:maxCont}, the concretizations of the children of $P$ require no process of sort $\PHsort(P)$.
      Furthermore, there exists $\zeta \in \BS(P)$ so that $(P, \aZ) \in \Bsol$.
      We build recursively a scenario $\delta$. Let $m = |\indexes{\zeta}|$.
      \begin{itemize}
        \item[*] \towrite{construire récurvivement ? Distinguer les cas $b_j$ dans $\components$ ou $\cs$ (stabilité requise et obtenue)}
        For $n \in \indexes{\zeta}$, let $s_n = s \PHplay \delta_1 \PHplay \delta'_1 \PHplay \delta''_1 \PHplay \zeta_1 \PHplay \dots \PHplay \delta_{n-1} \PHplay \delta'_{n-1} \PHplay \delta''_{n-1} \PHplay \zeta_{n-1}$ (or $s_1 = s$),
          and $\zeta_n = \PHhit{b_i}{a_j}{a_k}$.
          By hypothesis, $\exists \delta_n \in \muconcr_{s_n}(\PHobj{\any}{b_i})$.
          If $a \in \cs$, then $\prio(\zeta_n) = 1$ and $\zeta_n$ can be played in $s_n \PHplay \delta_n$ (\ie $\delta'_n = \delta''_n = \varepsilon$).
          If $a \notin \cs$ and $b \in \cs$, then there exists a scenario $\delta'_n$ so that $b_i \in \pfp_{s_n \PHplay \delta_n \PHplay \delta'_n}(b)$ (because of the $\Bpriocont$ relation).
          Then, by \pref{th:hcscomp}, there is a scenario $\delta'' \in \restriction{\Sce}{1}$ so that $\zeta_n$ is playable in $s_n \PHplay \delta_n \PHplay \delta'_n \PHplay \delta''_n$.
          Finally, if $a,b \in \components$, then by \pref{th:hcompcomp}, there is a scenario $\delta'' \in \restriction{\Sce}{1}$ so that $\zeta_n$ is playable in $s_n \PHplay \delta_n \PHplay \delta''_n$ (\ie $\delta'_n = \varepsilon$).
      \end{itemize}
      Thus, $\delta = \delta_m \in \muconcr_s(P)$. % and $\ceil(\delta) \subseteq max\ctx$.
  \end{itemize}
\end{itemize}

>>>>

\begin{definition}
  \label{def:aS}
  The abstract structure $\cwB=(\Breq,\Bsol,\Brsp,\Bcont)$ is defined as
  $\cwB = \sfp{\aB^\w_\ctx}{\myB}{\aB^\w_{\ctx \Cap \allprocs_\ctx(\myB)}}$,\\
  with $\myB=(\myreq,\mysol,\myrsp,\mycont)$:
  \begin{align*}
    \myreq &= \{ (a_i,\PHobjp{a}{j}{i}) \in \PHproc \times \Obj \mid
      a \in \components \wedge a_j \in \PHget{\ctx}{a} \\ % \vee a_j \in \pfpprocs_\ctx(a) \\
      & \qquad \wedge ((\exists (P,ps) \in \mysol \vee \exists (b_j, ps) \in \myrsp), a_i \in ps \\
      & \qquad\qquad \vee \exists n \in \indexes{\w}, \PHbounce(\w_n)=a_i) \}
    \\
    \mysol &\subseteq \{ (P,ps) \in \Obj \times \powerset(\PHproc) \mid
            \exists (a_i, P) \in \myreq \wedge ps \in \aBS(P) \\
      & \qquad\qquad \vee \exists (Q, P) \in \mycont \wedge ps \in \aBS(P) \}
    \\
    \myrsp &= \{ (a_i,ps) \in \PHproc \times \powerset(\PHproc) \mid a \in \cs \\
      & \qquad \wedge (\exists (P,ps') \in \mysol \vee \exists (b_j,ps') \in \myrsp), \\
      & \qquad\qquad a_i \in ps' \wedge \csState(a_i) \in ps \}
    \\
    \mycont & = \{ (P, \PHobj{q}{\PHbounce(P)}) \in \Obj \times \Obj \mid
      \exists (P, ps) \in \mysol \\
      & \qquad\qquad \wedge q \in \gCont_\ctx(\PHsort(P),P) \}
  \end{align*}
\end{definition}

\begin{definition}[Coherent solution]
\label{def:coherent}
  A cooperative sort requisite $(a_i, ps)$ in $\myrsp$ is said coherent iff
  it has no successor $(b_k,\PHobjp{b}{j}{k})$ in $\myreq$ so that there exists $b_n \in ps$, $b_k \neq b_n$.
\end{definition}

\begin{theorem}[Approximation inf.]
\label{th:approxinf}
  If the graph $\cwB$ contains no cycle,
  all objectives have at least one solution
  and all cooperative sort requisites are coherent,
  then $\uconcr(\w) \neq \emptyset$.
\end{theorem}

@@

Let $(a_i, ps) \in \Brsp$ be a cooperative sort requirement whom all children are concretizable.
We label all processes of $ps$ by an integer: $ps = \{ p_n \}_{n \in \indexes{ps}}$.
Let us prove that for all $n \in \indexes{ps}$, there exists a scenario $\delta_n$ so that:
$\forall i \in \segm{1}{n}, \PHget{(s \PHplay \delta_n)}{\PHsort(p_i)} = p_i$.
\begin{itemize}
  \item It is straightforward for $\delta_0 = \varepsilon$.
  \item Let us suppose such $\delta_n$ exists. \todo{Il faut être plus précis pour prendre en compte les coopérations en cascade}
    By hypothesis, there exists $\delta' \in \muconcr_{s \PHplay \delta_n}(\PHobj{\any}{p_n})$.
%    so that: $\PHget{(s \PHplay \delta_n \PHplay \delta')}{\PHsort(p_{n+1})} = p_{n+1}$.
    By hypothesis, $(a_i, ps)$ is coherent (\pref{def:coherent}), which means that solving $p_{n+1}$ does not disturb the processes of $ps$.
    Therefore, $\forall i \in \segm{1}{n+1}$, $\PHget{(s \PHplay \delta_n \PHplay \delta')}{\PHsort(p_{i})} = p_{i}$.
    Finally, by theorem \todo{théorème mise à jour}, there exists a scenario $\delta'' \in \restriction{\Sce}{1}$
    that can be played in $s \PHplay \delta_n \PHplay \delta'$
    and so that, if we denote $\delta_{n+1} = \delta_n \PHplay \delta' \PHplay \delta''$,
    we have: $\update(s \PHplay \delta_n \PHplay \delta') = s \PHplay \delta_{n+1}$ (\todo{théorème stabilité}), and the same property about processes.
\end{itemize}
Therefore, $\delta = \delta_{\indexes{ps}}$ exists, and given its properties, we have: $\PHget{(s \PHplay \delta)}{a} = a_i$
and $\update(s \PHplay \delta) = s \PHplay \delta$.

@@

  \item Suppose all children objectives of $P$ are concretizable.
  \begin{itemize}
    \item If $\exists (P, Q) \in \Bcont$, then by hypothesis,
      $\muconcr_{s}(\obj{\PHtarget(P)}{\PHtarget(Q)} \concat Q) \neq \emptyset$, thus
      $\muconcr_{s}(P) \neq \emptyset$.
    \item Else, by \pref{def:maxCont}, the concretizations of the children of $P$ require no process of sort $\PHsort(P)$.
      Furthermore, there exists $\zeta \in \BS(P)$ so that $(P, \aZ) \in \Bsol$.
      We show by induction that for all $n \in \indexes{\zeta}$, there is a scenario $\delta_n$ so that $\PHget{(s \PHplay \delta_n)}{\PHsort(P)} = \PHbounce(\zeta_n)$.
%      We build recursively a scenario $\delta$. Let $m = |\indexes{\zeta}|$.
      \begin{itemize}
        \item[*] \towrite{construire récurvivement ? Distinguer les cas $b_j$ dans $\components$ ou $\cs$ (stabilité requise et obtenue)}
        Suppose that $\delta_n$ exists and let $\zeta_n = \PHhit{b_i}{a_j}{a_k}$.
        By hypothesis there exists $\delta' \in \muconcr_{s \PHplay \delta_n}(\PHobj{\any}{b_i})$ with $\PHsort(P) \notin \PHsort(\delta')$ by \pref{def:maxCont}.
        By \todo{théorème mise à jour} there exists $\delta'' \in \restriction{\Sce}{1}$ so that $\update(s \PHplay \delta') = s \PHplay \delta' \PHplay \delta''$.
        Furthermore, $\PHget{(s \PHplay \delta' \PHplay \delta'')}{b} = b_j$ \todo{car $b \in \components$ ou $b$ est stable}.
        Therefore, $\delta_{n+1} = \delta_n \PHplay \delta' \PHplay \delta'' \PHplay \zeta_n$.
      \end{itemize}
      Thus, $\delta_{\indexes{\zeta}} \in \muconcr_s(P)$. % and $\ceil(\delta) \subseteq max\ctx$.
  \end{itemize}
\end{itemize}
